---
title: "p8105_hw6_kg3006"
author: "Kevin Guzman"
date: "2025-11-25"
output: github_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Loading appropriate libraries 
library(tidyverse)
library(purrr)
library(broom)
library(p8105.datasets)
library(gridExtra)
```

## Problem 1

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#Loading library from github website 
homicides <- read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")

#Creating city, state variable using "city", "state" 
homicides <- homicides %>%
  mutate(city_state = paste(city, state, sep = ", ")) %>% 
  #Creating binary variable (solved v. not solved) from disposition column
  mutate(solved = ifelse(disposition == "Closed by arrest", 1, 0)) %>% 
  #Omiting city_states Dallas, Tx, Phoenix, AZ, and Kansas City, MO for not reporting race 
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) %>% 
  #Omitting Tulsa, AL (entery mistake) 
  filter(city_state != "Tulsa, AL") %>% 
  #Only keeping those who are black or white race in victim_race
  filter(victim_race==c("Black", "White")) %>% 
  #Changing age to nmeric 
  mutate(victim_age = as.numeric(victim_age)) 
```

### Comparing solving crimes in males v. females in Baltimore, MD
 
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#Baltiomore Analysis 
  #Filtering for Baltimore, MD to new dataset
  baltimore_homicides <- homicides %>%
    filter(city_state == "Baltimore, MD")
  
  #Logistic regression model predicting whether homicide is solved based on victim age, sex, and race as predictors 
  baltimore_glm <- glm(solved ~ victim_age + victim_sex + victim_race, 
                      data = baltimore_homicides,
                      family=binomial)
  #Saving as an R object 
  saveRDS(baltimore_glm, "Output/baltimore_glm.rds")
  
  #Getting adjusted odds ratios 
  baltimore_glm %>% 
    broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

In Balitmore, MD, the adjusted odds of a homicide being solved for Males are `r round( (baltimore_glm %>% broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>% filter(term == "victim_sexMale"))$estimate, 2)` (95% CI: `r round( (baltimore_glm %>% broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>% filter(term == "victim_sexMale"))$conf.low, 2)` - `r round( (baltimore_glm %>% broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>% filter(term == "victim_sexMale"))$conf.high, 2)`) times the odds for female victime, holding all other variables constant.

### All cities 

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.width=8, fig.height=10}
# Making odds ratio for each city using purr and piplein 
  city_odds_ratio <- homicides %>% 
    #Group by city_state 
    group_by(city_state) %>% 
    #Nest data
    nest() %>% 
    #Creating function to run logistic regression for each city
    mutate(
      model = map(data, ~ glm(solved ~ victim_age + victim_sex + victim_race,
                              data = .x,
                              family = binomial)),
    #Getting adjusted odds ratios for each city
    tidy = map(model, ~ broom::tidy(.x, exponentiate = TRUE, conf.int = TRUE))) %>% 
    select(city_state, tidy) %>% 
    unnest(cols = c(tidy)) %>% 
    #Filtering for victim_sexMale term only
    filter(term == "victim_sexMale") %>% 
    #Only keep estimates and OR high and low 
    select(city_state, estimate, conf.low, conf.high) %>%
    #Arranging in descending order of odds ratio
    arrange(desc(estimate)) 

    #Creating plot that shows estimate OR and CI for each city 
    plot1 <- city_odds_ratio %>% 
      ggplot(aes(x = reorder(city_state, estimate), y = estimate)) +
      # Vertical reference line at OR = 1
      geom_hline(yintercept = 1, linetype = "dashed", color = "gray40", linewidth = 0.7) +
      geom_point(size = 3, color = "black") +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = 0.15,
                color = "gray20",
                linewidth = 0.6) +
      coord_flip() +
      labs(
        title = "Homicide Solving by Victim Sex (Male vs. Females)",
        x = "City, State",
        y = "Adjusted Odds Ratio (Male vs. Female)"
      )  +
      theme_minimal(base_size = 14) +
      theme(
        plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(size = 12, color = "gray25"),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(color = "gray90"),
        plot.margin = margin(10, 20, 10, 10)
      ) + 
      #Adding caption 
      labs(caption = "Adjusted odds ratio of the probability of a homicide being solved based on the victim's sex")
    
    plot1

```

Rates of homicides being solved differ by cities. The city with the highest adjusted odds of a homicide being solved when the victim is male compared to female is `r city_odds_ratio$city_state[which.max(city_odds_ratio$estimate)]` with an adjusted odds ratio of `r round(max(city_odds_ratio$estimate), 2)` (95% CI: `r round(city_odds_ratio$conf.low[which.max(city_odds_ratio$estimate)], 2)` - `r round(city_odds_ratio$conf.high[which.max(city_odds_ratio$estimate)], 2)`). The city with the lowest adjusted odds of a homicide being solved when the victim is male compared to female is `r city_odds_ratio$city_state[which.min(city_odds_ratio$estimate)]` with an adjusted odds ratio of `r round(min(city_odds_ratio$estimate), 2)` (95% CI: `r round(city_odds_ratio$conf.low[which.min(city_odds_ratio$estimate)], 2)` - `r round(city_odds_ratio$conf.high[which.min(city_odds_ratio$estimate)], 2)`).

## Problem 2 

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.width=8, fig.height=10}
#Loading library from p8105.datasets 
  data("weather_df")

#Viewing 
  head(weather_df)

#5000 bootstraps of model 
  #setting seed
  set.seed(123)
  
  #making function for bootstrap 
  bootstrap_function <- function(data, indices) {
    d <- data[indices, ] # Resample the data
    model <- lm(tmax ~ tmin + prcp, data = d)
    #Extract r^2
    r2 <- glance(model)$r.squared
    
    #extract coefficients
    coefs <- tidy(model)
    b1 <- coefs$estimate[coefs$term == "tmin"]
    b2 <- coefs$estimate[coefs$term == "prcp"]
    
    tibble(r_squared = r2,
           beta1_beta2 = b1*b2)   
  }
  
  #Running 5000 boostrap samples 
  bootstrap_results <- map_dfr(1:5000, 
                               ~ bootstrap_function(weather_df, sample(1:nrow(weather_df), replace = TRUE)))
  
  #Plotting results 
  p1 <- ggplot(bootstrap_results, aes(x = r_squared)) +
  geom_histogram(color = "white", fill = "gray40", bins = 40) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Bootstrap Distribution of R-squared",
    x = expression(hat(R)^2),
    y = "Frequency"
  ) + 
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black")
  )
  
  p2 <- ggplot(bootstrap_results, aes(x = beta1_beta2)) +
  geom_histogram(color = "white", fill = "gray40", bins = 40) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Bootstrap Distribution of β1 × β2",
    x = expression(hat(beta)[1] * hat(beta)[2]),
    y = "Frequency"
  ) + 
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black")
  )
  
  
  #Combine figures
  grid.arrange(p1, p2, ncol = 1)
```


The bootstrap distribution of R-squared shows that the R-squared values fall between `r round(min(bootstrap_results$r_squared), 3)` and `r round(max(bootstrap_results$r_squared), 3)`. The mean R-squared was `r round(mean(bootstrap_results$r_squared), 3)` (95% CI: `r round(quantile(bootstrap_results$r_squared, c(0.025, 0.975)),3)`) indicating that the model explains a substantial portion of the variance in maximum temperature (tmax).  The bootstrap distribution of β̂₁β̂₂ ranged from`r round(min(bootstrap_results$beta1_beta2), 3)` to `r round(max(bootstrap_results$beta1_beta2), 3)`. The mean value is `r round(mean(bootstrap_results$beta1_beta2), 3)` (95% CI: `r round(quantile(bootstrap_results$beta1_beta2, c(0.025, 0.975)),3)`). 


## Problem 3

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#Loading CSV dataset from Files
  birthweight <- read_csv("Files/birthweight.csv") %>% 
  #Sex as factor 1 = Male, 2 = Female
  mutate(babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female"))) %>% 
  #frace :father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
  mutate(frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), 
                        labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown"))) %>% 
  #malform : presence of any malformation (0 = absent, 1 = present)
  mutate(malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present"))) %>% 
  #mrace: (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
  mutate(mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), 
                        labels = c("White", "Black", "Asian", "Puerto Rican", "Other"))) %>% 
  rename(
    baby_sex = babysex,
    baby_head_cm = bhead,
    baby_length_cm = blength,
    baby_weight_g = bwt,
    mothers_delivery_weight_lb = delwt,
    family_income = fincome,
    fathers_race = frace,
    gestation_week = gaweeks,
    malformation = malform,
    mothers_age_menarche = menarche,
    mothers_height_inches = mheight,
    mothers_delivery_age = momage, 
    mothers_race = mrace, 
    prior_births = parity,
    prior_low_birthweight = pnumlbw,
    prior_small_gestation = pnumsga,
    mothers_pre_preg_BMI = ppbmi,
    mothers_pre_preg_wegith_lb = ppwt,
    cigarette_preganancy = smoken,
    preganancy_wt_gain_lb = wtgain
  ) %>% 
  #change pounds to kg
    mutate(mothers_delivery_weight_kg = mothers_delivery_weight_lb * 0.453592,
           mothers_pre_preg_weight_kg = mothers_pre_preg_wegith_lb * 0.453592,
           pregnancy_wt_gain_kg = preganancy_wt_gain_lb * 0.453592) %>%
  #change inches to cm
    mutate(mothers_height_cm = mothers_height_inches * 2.54) %>%
  #change g to kg 
    mutate(baby_weight_kg = baby_weight_g / 1000) %>%
  #Remove old variables
    select(-c(mothers_delivery_weight_lb, preganancy_wt_gain_lb, mothers_pre_preg_wegith_lb, mothers_height_inches, baby_weight_g))


  
    



```

